Characterizing Optimizations to Memory Access Patterns using Architecture-Independent Program Features

### 2020-01-19 paper/paper.tex
name the PSL metric earlier in relevant section

	AIWC's metrics strongly reward optimizations that tend to localize memory accesses. Local memory buffers are typically smaller than global memory arrays when programming for GPUs due to hardware limitations on sizes of shared memory \cite{cudamanual}. However, the metrics currently measured by AIWC do not have a direct causal relation to code patterns that optimize memory accesses on GPUs. The proposed relative local memory usage metric is the first to correspond to a recommended optimization strategy of using fast on-chip shared memory. Further, we find that all current AIWC metrics do not measure any sizeable difference between \texttt{coalescedAB}, \texttt{coalescedABT} and \texttt{alignedABT} codes. We address the root cause of this -- a lack of temporal information about memory accesses, in the next section.%TODO READ THIS SECTION

@@

	We propose a new architecture-independent metric to measure memory access patterns of parallel programs.
	The proposed metric is inspired by CuMAPz' direct approaches to measuring optimization specific characteristics of CUDA codes.

@@

	\textit{Calculate}: for each timestamp $t = t_0 .. t_{last}$, calculate the n-bits dropped entropy of memory addresses accessed by all threads in a work group within the timestamp $t$. Here $n$ can range between 0 and 10 as was the case for $LMAE$.

	\textit{Summarize}: average the collected entropy values across all the timestamps to calculate the parallel spatial locality metrics for one thread group. We then average the n-bits dropped entropy summaries across thread groups to obtain the \textit{n-bits dropped parallel spatial locality metrics} for the kernel's execution.

@@

%	With $E_{n,t}$ the n-bits dropped entropy at timestamp $t$.

	The proposed metric is a parallel computing analogue for MICA's data stride metric that measures the distance between consecutive data accesses in a single-threaded environment. 
	In parallel programs, to accurately measure spatial locality of accesses, we must consider memory accesses performed by multiple threads in a close temporal scope. 
	The proposed metric calculates the locality of accesses in each time step of the program's execution and steeper reductions in $n$-bits dropped parallel spatial locality scores will be observed in programs that often access nearby memory addresses within the same timestamp.
	Such programs will perform better on GPUs, as they will make better use of both global memory access coalescing and shared memory bank structures.
	To a lesser extent, the proposed metric reflects performance-critical memory access patterns on CPUs, as pulling a single cache line from global memory into last-level cache may improve memory access times for all CPU cores.

>>>>

	AIWC's metrics strongly reward optimizations that tend to localize memory accesses. Local memory buffers are typically smaller than global memory arrays when programming for GPUs due to hardware limitations on sizes of shared memory \cite{cudamanual}. However, the metrics currently measured by AIWC do not have a direct causal relation to code patterns that optimize memory accesses on GPUs. The proposed relative local memory usage metric is the first to correspond to a recommended optimization strategy of using fast on-chip shared memory. Further, we find that all current AIWC metrics do not measure any sizeable difference between \texttt{coalescedAB}, \texttt{coalescedABT} and \texttt{alignedABT} codes.
	We address this by proposing another new metric for locality of memory accesses in the following section.

@@

	We propose a new architecture-independent metric, \emph{parallel spatial locality}, to measure memory access patterns of parallel programs.
	The proposed metric is inspired by CuMAPz' direct approaches to measuring optimization specific characteristics of CUDA codes.

@@

	\textit{Calculate}: for each timestamp $t = t_0 .. t_{last}$, calculate the $n$-bits-dropped entropy of memory addresses accessed by all threads in a work group within the timestamp $t$. Here $n$ can range between 0 and 10 as was the case for $LMAE$.

	\textit{Summarize}: average the collected entropy values across all the timestamps to calculate the parallel spatial locality metrics for one thread group. We then average the $n$-bits-dropped entropy summaries across thread groups to obtain the $n$-bits-dropped parallel spatial locality metrics for the kernel's execution.

@@

%	With $E_{n,t}$ the n-bits-dropped entropy at timestamp $t$.

	The proposed metric is a parallel computing analogue for MICA's data stride metric that measures the distance between consecutive data accesses in a single-threaded environment. 
	In parallel programs, to accurately measure spatial locality of accesses, we must consider memory accesses performed by multiple threads in a close temporal scope. 
	The proposed metric calculates the locality of accesses in each time step of the program's execution and steeper reductions in $n$-bits-dropped parallel spatial locality scores will be observed in programs that often access nearby memory addresses within the same timestamp.
	Such programs will perform better on GPUs, as they will make better use of both global memory access coalescing and shared memory bank structures.
	To a lesser extent, the proposed metric reflects performance-critical memory access patterns on CPUs, as pulling a single cache line from global memory into last-level cache may improve memory access times for all CPU cores.
